Profit Optimisation in Heterogeneous 2-link Parallel Games

### 2024-09-28 thesis/article.tex
Thesis: Adds in Pricing Games basics for fixed dist func case

% Commands for wrapping properly common expressions.
\newcommand{\indeq}[1]{\stackrel{\text{#1}}{=}}
\newcommand{\RightarrowArg}[1]{\stackrel{#1}{\Rightarrow}}
\newcommand{\LeftrightarrowArg}[1]{\stackrel{#1}{\Leftrightarrow}}
\newcommand{\NE}{\mathrm{N.E.}}
\newcommand{\as}{\mathrm{\alpha_s}}
\newcommand{\R}{\mathbb{R}}
\DeclareMathOperator*{\argmax}{arg\,max}
% \newcommand{\Exp}{\mathrm{Exp}}
% \newcommand{\Expect}{{\rm I\kern-.3em E}}

@@

>>>>

% Commands for wrapping properly common expressions.
\newcommand{\indeq}[1]{\stackrel{\text{#1}}{=}}
\newcommand{\RightarrowArg}[1]{\stackrel{#1}{\Rightarrow}}
\newcommand{\LeftrightarrowArg}[1]{\stackrel{#1}{\Leftrightarrow}}
\newcommand{\NE}{\mathrm{N.E.}}
\newcommand{\as}{\mathrm{\alpha_s}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Gm}{\mathcal{G}}
\DeclareMathOperator*{\argmax}{arg\,max}
% \newcommand{\Exp}{\mathrm{Exp}}
% \newcommand{\Expect}{{\rm I\kern-.3em E}}

@@

For the case where the distribution function of our model is fixed, i.e. $\alpha(p) = c$ for some $c > 0$, then the game becomes homogeneous again, simply with a different -common- sensitivity to tolls.
However, what interests us in this study is the relation between games for different $c > 0$, as it will prove crucial when analysing step distribution functions later.
We start with formalising our base.

\begin{lemma}
	\label{lemma:a_fixed_base}
	Let $\Gm_1 = ([2], \ell^{(1)}, c)$ a heterogeneous pricing game with distribution function $\alpha(p) = c$ for some $c > 0$ and $\Gm_2 = ([2], \ell^{(1)} / c, 1)$ a homogeneous pricing game with latencies $\ell^{(1)} / c$ for the same $c$.
	Then the two games are equivalent.
\end{lemma}

\begin{proof}
	It suffices to show that $x^{(1)}(t) = x^{(2)}(t), \; \forall t \in \R_+^2 \cup \{0\}$, where $x^{(1)}(t), x^{(2)}(t)$ the respective Nash Equilibrium for $t$ for $\Gm_1, \Gm_2$.
	Begin by noticing that while $\Gm_1$ is heterogeneous, all its players have the same money sensitivity, and thus at an equilibrium they will all see the two edges as the same.
	Therefore for all $t \in \R_+ \cup \{0\}$, $x^{(1)}(t)$ is the unique solution of the equation
	\[
		\ell_1^{(1)}(x_1^{(1)}(t)) + ct_1 = \ell_2^{(1)}(x_2^{(1)}(t)) + ct_2
	\]
	It is easy to check now that $x^{(2)}(t)$ is also a solution to the above equation, since in $\Gm_2$ it is the equivalent unique solution for
	\[
		\ell_1^{(1)}(x_1^{(2)}(t)) / c = \ell_2^{(1)}(x_2^{(2)}(t)) / c \Leftrightarrow \ell_1^{(1)}(x_1^{(2)}(t)) = \ell_2^{(1)}(x_2^{(2)}(t)) 
	\]
	Therefore $x^{(1)}(t) = x^{(2)}(t), \; \forall t \in \R_+^2 \cup \{0\}$, making $\Gm_1, \Gm_2$ equivalent.
\end{proof}
Intuitively, the above means that increasing the money sensitivity of the players in equivalent to analogously decreasing the traffic that flows create on all edges.
Unfortunately this correspondence  disappears for non-fixed distribution functions, since, as we saw in Chapter 2, we loose the $1-1$ relation between the latency functions and the distribution function due to heterogeneity.

Still, looking now at the relation between fixed distribution functions, Lemma \ref{lemma:a_fixed_base} gives us the following corollary.
\begin{corollary}
		Let $\Gm_1 = ([2], \ell, c_1)$ and $\Gm_2 = ([2], \ell, c_1)$ two heterogeneous pricing games with the same latency functions $\ell$ and distribution functions $\alpha(p) = c_1 > 0$ and $c_2 > 0$ respectively.
		Then the following are true
		\begin{enumerate}[$(i)$]
			\item $x^{(1)}(0) = x^{(2)}(0)$
			\item $x^{(1)}(t) = x^{(2)}\left(\frac{c_1}{c_2} \cdot t\right)$
			\item $\Pi_i^{(1)}(t_i, t_{-i}) = \frac{c_2}{c_1} \cdot \Pi_i^{(2)}\left(\frac{c_1}{c_2} \cdot t_i, \frac{c_1}{c_2} \cdot t_{-i}\right)$
			\item $B_i^{(1)}(t_{-i}) = \frac{c_2}{c_1} \cdot B_i^{(2)}\left(\frac{c_1}{c_2} \cdot t_{-i}\right)$
		\end{enumerate}
\end{corollary}

\begin{proof}
	$ $
	\begin{enumerate}[$(i)$]
		\item Obvious since for $t = 0$ the flow in both $\Gm_1, \Gm_2$ play a homogeneous game with common latencies $\ell$, which makes the two games identical and therefore $x^{(1)}(0) = x^{(2)}(0)$.
		\item With the same rational as in Lemma \ref{lemma:a_fixed_base}, looking at the respective cost equalities for the two games we see that starting from $\Gm_2$ for tolls $(c_1 / c_2)t$ we get
		\begin{flalign*}
			\ell_1&\left(x_1^{(2)}\left(\frac{c_1}{c_2} \cdot t\right)\right) + c_2 \cdot \left(\frac{c_1}{c_2} \cdot t_1\right) = \ell_2\left(x_2^{(2)}\left(\frac{c_1}{c_2} \cdot t\right)\right) + c_2 \cdot \left(\frac{c_1}{c_2} \cdot t_2\right) & \\
			&\Rightarrow \ell_1\left(x_1^{(2)}\left(\frac{c_1}{c_2} \cdot t\right)\right) + c_1 t_1 = \ell_2\left(x_2^{(2)}\left(\frac{c_1}{c_2} \cdot t\right)\right) + c_1 t_2 &
		\end{flalign*}
		with the last equation being the one for edge equality in $\Gm_1$ with a unique solution at $x^{(1)}(t)$, leading us to $x^{(1)}(t) = x^{(2)}\left(\frac{c_1}{c_2} \cdot t\right)$
		\item From $(ii)$ and the definition of $\Pi_i$ we get
		\begin{flalign*}
			\Pi_i^{(1)}(t_i, t_{-i}) &= x_i^{(1)}(t) \cdot t_i \stackrel{(ii)}{=} x_i^{(2)}\left(\frac{c_1}{c_2} \cdot t\right) \cdot t_i & \\
			&= \frac{c_2}{c_1} \cdot \Pi_i^{(2)}\left(\frac{c_1}{c_2} \cdot t_i, \frac{c_1}{c_2} \cdot t_{-i}\right) &
		\end{flalign*}
		\item From $(iii)$ and the definition of $B_i$ we get
		\begin{flalign*}
			B_i^{(1)}(t_{-i}) &= \argmax_{t_i > 0}\Pi_i^{(1)}(t_i, t_{-i}) \stackrel{(iii)}{=} \argmax_{t_i > 0}\frac{c_2}{c_1} \cdot \Pi_i^{(2)}\left(\frac{c_1}{c_2} \cdot t_i, \frac{c_1}{c_2} \cdot t_{-i}\right) & \\
			&= \frac{c_2}{c_1} \cdot B_i^{(2)}\left(\frac{c_1}{c_2} \cdot t_{-i}\right) &
		\end{flalign*}
	\end{enumerate}
\end{proof}

While the above properties seem straightforward, they are of little use to us as they are.
What we actually need to examine is how those two games react differently for an edge $i$ given a fixed $t_{-i}$.
